---
title: "Set up ED2 runs"
author: "Eric R. Scott"
date: "2022-10-13"
output: html_document
---
```{r setup}
library(tidyverse)
#need dev version of PEcAn.settings to strip comments
# remotes::install_github("pecanproject/pecan", subdir = "base/settings")
library(PEcAn.settings) 
```

Need to:

1. Copy over a template
2. Create directories
2. Edit template pecan.xml to include site ID, site name, paths to history files, etc.
3. Copy relevant files from Welsch to the HPC
4. Edit template workflow.R to point to edited pecan.xml.

To figure out how to automate all this, start by picking a site (literally at random) and setting up a run:

```{r}
new_site <- read_csv("data/mandifore_sites.csv")
set.seed(123)
(s <- slice_sample(new_sites, n = 1))
```

# 1. Copy over templates

Make directory

```{r}
wd <- file.path("MANDIFORE_runs", s$sitename)
dir.create(wd)
```

Copy templates

```{r}
file.copy(file.path("templates", "pecan_template.xml"), file.path(wd, "pecan.xml"))
file.copy(file.path("templates", "workflow_template.R"), file.path(wd, "workflow.R"))
```


# 2. Customize pecan.xml

I think this will be easiest to work with it in list format

```{r}
settings <- read.settings(file.path(wd, "pecan.xml"))
```

## Outdir

For now I'll put it inside the run folder.  Later move to `/data/`

```{r}
settings$outdir <- file.path(wd, "outdir")
dir.create(settings$outdir)
```


## Site name and ID

```{r}
settings$info$notes <- s$sitename

settings$run$site$id <- s$site_id
```

## MET data

```{r}
settings$run$site$met.start <- format(s$met_start_date, "%Y-%m-%d %H:%M:%S")
settings$run$site$met.end <- format(s$met_end_date, "%Y-%m-%d %H:%M:%S")
```

Need to edit ED_MET_DRIVER_HEADER to point to correct path

Change `/data/input/NARR_ED2_site_1-18168/` to `/data/sites/mandifore/NARR_ED2_site_1-18168/`

```{r}
#need to copy without permissions (copy.mode = FALSE) to a new directory to edit
file.copy(
  from = file.path("/data/input", s$met_filename),
  to = "/data/sites/mandifore",
  recursive = TRUE,
  copy.mode = FALSE
)

met_driver_path <- file.path("/data/sites/mandifore", s$met_filename, "ED_MET_DRIVER_HEADER")
met_driver <- readLines(met_driver_path) #read in
met_driver <- str_replace(met_driver, "/data/input/", "/data/sites/mandifore/") #fix path
write_lines(met_driver, met_driver_path)
```

## History files

```{r}
base <- "/data/sites/mandifore"

settings$run$inputs$met <- file.path(base, s$met_filename, "ED_MET_DRIVER_HEADER")

files <- list.files(file.path("/data/input/", s$cohort_filename))

settings$run$inputs$pss <-
  file.path(base, s$cohort_filename, files[str_detect(files, "\\.pss$")])
settings$run$inputs$css <-
  file.path(base, s$cohort_filename, files[str_detect(files, "\\.css$")])
settings$run$inputs$site <-
  file.path(base, s$cohort_filename, files[str_detect(files, "\\.site$")])
```

## PFTs

I think I need to add PFTs to `settings` from the .css file

```{r}
css <-
  read_table2(list.files(
    file.path("/data/input/", s$cohort_filename),
    pattern = ".css$",
    full.names = TRUE
  ))
pfts <- tibble(ED = unique(css$pft))

#TODO: this matches multiple PEcAn PFTs per ED PFT number.  So this is likely not going to work without some additional data or by manually specifying.  For now as proof of concept I'll just pick the first match
pfts <- left_join(pfts, PEcAn.ED2::pftmapping, by = "ED") %>% group_by(ED) %>% slice(1)

for (i in 1:nrow(pfts)) {
  new_pft <- list(
    pft = list(
      name = pfts$PEcAn[i],
      ed2_pft_number = as.character(pfts$ED[i])
    )
  )
  settings$pfts <- append(settings$pfts, new_pft)
}


settings$pfts
```



# 3. check settings

```{r}
settings <- prepare.settings(settings)
write.settings(settings, outputfile = "settings_checked.xml")
```



# 4. Customize workflow.R

```{r}
workflow <- readLines(file.path(wd, "workflow.R"))
checked_settings_path <- file.path("ED2", "MANDIFORE_runs", s$sitename, "outdir", "settings_checked.xml")
repl <- paste0('inputfile <- \"', checked_settings_path, '\"')
workflow <- str_replace(workflow, 'inputfile <- .*', replacement = repl)
writeLines(workflow, file.path(wd, "workflow.R"))
```

# 5. Copy files over to HPC

The weather and history files need to be on the HPC

```{r}
PEcAn.remote::remote.copy.to(
  settings$host,
  src = file.path("/data/sites/mandifore/", s$met_filename),
  dst = "/groups/dlebauer/data/sites/mandifore"
)
PEcAn.remote::remote.copy.to(
  settings$host,
  src = file.path("/data/input", s$cohort_filename),
  dst = "/groups/dlebauer/data/sites/mandifore"
)
```