---
title: "Designing patches representing the ecosystems of the Southeast US"
author: Eric R. Scott
output: html_document
date: '2023-01-11'
---

```{r}
library(spatstat.data)
library(spatstat.geom)
library(tidyverse)
```

# Ecoregions

[Southern Coastal Plain](https://bplant.org/region/134): 
 - longleaf pine flatwoods and savannas
 - richer forests with slash pine, pond pine, pond cypres, american sweetgum, southern magnolia, laurel oak, white oak, american beech (i.e. mix of evergreen, evergreen hardwood, and deciduous hardwood PFTs)
 - Floodplains with bald cypres, pond cypres, water tupelo, sweetgum, green ash, water hickory ("temperate.Hydric" PFT??)
 - Florida scrub
 - monoculture pine plantations

[Southeastern plains](https://bplant.org/region/123): 
- more hardwoods and more closed-canopy than southern coastal plain
- pine forests with longleaf pine and loblolly pine
- Mixed oak-hickory-pine forests
- southeastern mixed forest with mix of braodleaf evergreens, deciduous evergreens (SIC), and pines
- Floodplains had deciduous forests
- Cypress swamps

[Piedmont](https://bplant.org/region/102):
- Different bedrock / soil than other regions
- oak-pine-hickory forest with various Carya species, shortleaf pine, loblolly pine, white oak, post oak
- Agriculture, old-field, and early successional forest


Can probably summarize this as:
1. pine forest
2. oak-hickory-pine forest (deciduous hardwood and pine)
3. mixed forest (pine, deciduous hardwood, evergreen hardwood)
4. prairie (c4 and c3 grass, maybe forb)
5. old field (like above but with more forbs?) or early successional forest (like prairie but with small trees)

## Pine Forest

There's a longleaf pine dataset in the `spatstat.data` package that we can use.  It's 584 Longleaf pines in a 200m x 200m region in southern Georgia.  Data collected by Platt, Evans and Rathbun (1988).

```{r}
plot(longleaf)
longleaf_df <- as.data.frame(longleaf) |> 
  rename(dbh = marks)
```

Data needs:
- number of cohorts (bins)
- dbh for each cohort
- density (stem/m^2) for each cohort

```{r}
ggplot(longleaf_df, aes(dbh)) + geom_histogram()
```

Looks like we could use 2 or 3 cohorts:

Saplings (dbh 1--20 ish), mature trees (dbh 20-50 ish), old trees (50-70 ish)

```{r}
longleaf_df |> 
  mutate(group = case_when(
    dbh < 20 ~ "sapling",
    dbh >=20 & dbh < 50 ~ "mature tree",
    dbh >=50 ~ "old growth"
  )) |> 
  group_by(group) |> 
  summarize(
    n = n(),
    median_dbh = median(dbh),
    density = n/(200*200),
  )

```

### PFTs

- temperate.Southern_Pine (7)

## Oak-Hickory-Pine forest
https://www.ncwildlife.org/Portals/0/Conserving/documents/Piedmont/P_Oak_forest_mixed_hardwoods_pine.pdf?ver=2011-08-15-161006-140
"Usually these forests are uneven‐aged with old tress occasionally present."

### PFTs



## Mixed Forest

## Prairie

## Old field / early successional

