---
title: "MSR Dec 2022"
output: html_document
date: '2022-12-13'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(here)
library(tidyverse)
library(PEcAn.ED2)
library(tidync)
library(lubridate)
library(units)
# library(arrow)
library(furrr)
future::plan(future.callr::callr, workers = 5)
```

Get paths to all the runs.


```{r}
# base_dir <- here("completed_runs") #for testing
base_dir <- here("MANDIFORE_runs")
sites <- list.dirs(base_dir, recursive = FALSE)
head(sites)
#for testing purposes
# sites <- sites[c(2,5:8)]
```

Not all of the runs completed and had successful conversion to .nc files.  I'll do this manually, I guess

```{r}
# library(furrr)
# future::plan(future.callr::callr, workers = 5)
# 
# site <- sites[5]
# ens_dirs <- 
#     list.files(file.path(site, "outdir", "out"),
#                pattern = "ENS-", full.names = TRUE)
# settings <-  PEcAn.settings::read.settings(here(site, "outdir", "settings_checked.xml"))
# #which runs don't have .nc files
# dirs_to_do <- ens_dirs[map_lgl(ens_dirs, ~length(list.files(.x, pattern = ".nc$")) == 0)]
# # dirs_to_do <- unique(c(dirs_to_do, ))
# # dirs_to_do <- dirs_to_do[c(-1, -6)]
# dirs_to_do <- dirs_to_do[c(-4, -5, -11)]
# #run model2netcdf.ED2() in parallel
# oldlevel <- PEcAn.logger::logger.setLevel("info")
# safe_model2netcdf <- safely(PEcAn.ED2::model2netcdf.ED2)
# future_map(dirs_to_do, ~{
#   pftmapping <- pftmapping #workaround, fixed in PEcAn pr
#   safe_model2netcdf(
#     outdir = .x,
#     settings = settings,
#     process_partial = TRUE #new arg
#   )
# }, .options = furrr_options(seed = TRUE))
# PEcAn.logger::logger.setLevel(oldlevel)
```


Get info about the run for plotting

```{r}
get_pft_names <- function(site) {
  settings_path <- 
    list.files(site, "settings_checked.xml", recursive = TRUE, full.names = TRUE)
  if(length(settings_path)==0) {
    warning("No settings file exists! Returning 0x0 tibble")
    return(tibble())
  }
  settings <- PEcAn.settings::read.settings(settings_path)
  x <- PEcAn.ED2:::extract_pfts(settings$pfts)
  pft_names <-
    tibble(pft = x, pft_name = names(x)) |> 
    mutate(site = basename(site)) |> 
    select(site, everything())
  return(pft_names)
}
# get_pft_names(sites[[3]])

pft_names <- map_df(sites, get_pft_names)
```

Get paths for output files and read in all the .nc files

```{r}
# nc_file <- here(base_dir, "MANDIFORE-PNW-4538", "outdir", "out", "ENS-00006-1000020684", "2003.nc")
  tidync(nc_file) |>
          activate("D0,D1,D5,D6") |> #these are the dims for variables that are separated by PFT
          hyper_tibble() |>
          mutate(date = make_date(year, 1, 1) + days(dtime))
```

```{r}
extract_nc <- function(nc_file) {
  nc <- nc_open(nc_file)
  pft <- ncvar_get(nc, varid = "PFT")
  dtime <- ncvar_get(nc, varid = "dtime")
  lat <- ncvar_get(nc, varid = "lat")
  lon <- ncvar_get(nc, varid = "lon")
  vars <- list(
    "AGB_PFT"    = ncvar_get(nc, varid = "AGB_PFT"    ),
    "BSEEDS"     = ncvar_get(nc, varid = "BSEEDS"     ),
    "DBH"        = ncvar_get(nc, varid = "DBH"        ),
    "DDBH"       = ncvar_get(nc, varid = "DDBH"       ),
    "NPP_PFT"    = ncvar_get(nc, varid = "NPP_PFT"    ),
    "TRANSP_PFT" = ncvar_get(nc, varid = "TRANSP_PFT" ),
    "DENS"       = ncvar_get(nc, varid = "DENS"       )
  )
  
  df <- 
    vars |> 
    map_df(~{
      colnames(.x) <- pft
      as_tibble(.x) |> 
        add_column(dtime)
    }, .id = "varid") |> 
    pivot_longer(c(-varid, -dtime),
                 names_to = "pft") |> 
    pivot_wider(names_from = varid, values_from = value) |> 
    add_column(lat, lon)
  nc_close(nc)
  df
}
```



```{r}
# site <- sites[[1]] #for debug
gather_data <- function(site) {
  ens_dirs <- 
    list.files(file.path(site, "outdir", "out"),
               pattern = "ENS-", full.names = TRUE)
  if(length(ens_dirs) == 0) {
    warning("No output found! Returning 0x0 tibble")
    return(tibble())
  }
  pb <- progress::progress_bar$new(total = length(ens_dirs))
  # .x <- ens_dirs[[6]] #for debug
  df_raw <- 
    map_dfr(ens_dirs, function(.x) {
      cat("processing ", .x)
      pb$tick()
      nc_files <- list.files(.x, pattern = "*.nc$", full.names = TRUE)
      if(length(nc_files) == 0 ) {
        warning("No .nc files found! Returning 0x0 tibble")
        return(tibble())
      }
      #extract and combine data from each year.nc file
      map_dfr(nc_files, function(.y) {
      # for (i in seq_along(nc_files)) {
        year <- stringr::str_remove(basename(.y), "\\.nc")
        tidync(.y) |>
          activate("D0,D1,D5,D6") |> #these are the dims for variables that are separated by PFT
          hyper_tibble() |>
          mutate(date = make_date(year, 1, 1) + days(dtime))
      # }
      })
    }, .id = "ensemble") |> 
    mutate(site = basename(site)) |> 
    select(site, ensemble, everything())
  return(df_raw)
}

# df_raw <- gather_data(site)
df_raw <- map_dfr(sites, gather_data)
```

Save raw data

```{r}
write_csv(df_raw, here("MSRs", "2022-12-15_data-raw.csv"))
```


Set units with the `units` package

```{r}
tidy_data <- function(df_raw) {
  df <- 
    df_raw |> 
    #set units
    mutate(
      AGB_PFT = set_units(AGB_PFT, "kg m-2"),
      NPP_PFT = set_units(NPP_PFT, "kg m-2 s-1"),
      TRANSP_PFT = set_units(TRANSP_PFT, "kg m-2 s-1"), #TODO check this!
      DENS = set_units(DENS, "1/m^2") #TODO not sure how to do stems/m2
    ) |>
    #calculate variables
    mutate(
      WUE_PFT = as.numeric(NPP_PFT / TRANSP_PFT)
    ) |> 
    mutate(
      WUE_PFT = case_when(
        as.numeric(NPP_PFT) <= 0 ~ 0,
        !is.finite(WUE_PFT) & as.numeric(NPP_PFT) > 0 ~ NA_real_,
        TRUE ~ WUE_PFT
      )
    ) |> 
    #convert units
    mutate(
      NPP_PFT = set_units(NPP_PFT, "kg m-2 day-1"),
      TRANSP_PFT = set_units(TRANSP_PFT, "kg m-2 day-1")
    ) |> 
    #first date is wonky, let's just remove it
    filter(date != min(date))
  df
}
df <- tidy_data(df_raw)
df <- left_join(df, pft_names) |> mutate(pft = as.factor(pft))
df
```

Summarize for plotting

```{r}
df_summary <- df |> 
  group_by(site, pft, pft_name, date) |> 
  summarize(
    across(
      c(NPP_PFT, AGB_PFT, DENS, TRANSP_PFT, WUE_PFT),
      .fns = list(
        mean = ~ mean(.x, na.rm = TRUE),
        lower = ~ quantile(.x, 0.25, na.rm = TRUE),
        upper = ~ quantile(.x, 0.75, na.rm = TRUE)
      )
    ),
    lat = first(lat),
    lon = first(lon)
  )
```

# Map

```{r}
library(maps)
US <- map_data("state")
sitemap <- 
  ggplot() + 
  geom_polygon(
    data = US,
    aes(x = long, y = lat, group = group),
    color = "black",
    fill =  "lightblue"
  ) +
  geom_point(
    data = df_summary,
    aes(x = lon, y = lat),
    fill = "red",
    color = "black",
    shape = 21,
    size  = 3
  ) +
  coord_map() +
  theme_void() +
  labs(title = "Site locations for test ED2 simulations") +
  theme(plot.title = element_text(hjust = 0.5))
ggsave(here("MSRs", "2022-12-15_map.png"), sitemap)
```


# Make plots

```{r}
source(here("R/shift_legend.R"))
```

Variables to plot:

- Density
- AGB
- Water use efficiency
- NPP


We want to highlight the progress in the following areas:
- More sites
- Longer runs
- Realistic communities

But we also want to make sure *Setaria* is visible.  This might require faceting by PFT with free Y axes or maybe making inset plots for *Setaria* using `ggforce`

```{r}
make_plot <- function(df_summary, y, ymin, ymax) {
  ggplot(df_summary, aes(x = date, color = pft_name, fill = pft_name)) +
    #uncomment to plot ensembles
    # geom_line(data = df, aes(y = AGB_PFT, group = ensemble), alpha = 0.4) +
    geom_line(aes(y = {{y}}), size = 0.7) +
    geom_ribbon(aes(ymin = {{ymin}}, ymax = {{ymax}}),
                color = NA, alpha = 0.4) +
    scale_color_brewer(type = "qual", palette = "Dark2", 
                       aesthetics = c("fill", "color")) +
    scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
    labs(
      color = "PFT",
      fill = "PFT",
      # y = "Water Use Efficiency",
      x = "Date"
    ) +
    theme_bw() +
    facet_wrap(~site, scales = "free_y")
}
```


```{r}
wue_plot <- 
  make_plot(df_summary, y = WUE_PFT_mean, ymin = WUE_PFT_lower, ymax = WUE_PFT_upper) + 
  labs(y = "Water Use Efficiency")

shift_legend(wue_plot) |> plot()
```


```{r}
agb_plot <- 
  make_plot(df_summary, y = AGB_PFT_mean, ymin = AGB_PFT_lower, ymax = AGB_PFT_upper) + 
  labs(y = "Aboveground Biomass")

shift_legend(agb_plot) |> plot()
```

Zoom in on setaria for 9095

```{r}
ggplot(df_summary, aes(x = date, color = pft_name, fill = pft_name)) +
    geom_line(aes(y = AGB_PFT_mean), size = 0.7) +
    geom_ribbon(aes(ymin = AGB_PFT_lower, ymax = {{ymax}}),
                color = NA, alpha = 0.4) +
    scale_color_brewer(type = "qual", palette = "Dark2", 
                       aesthetics = c("fill", "color")) +
    scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
    labs(
      color = "PFT",
      fill = "PFT",
      # y = "Water Use Efficiency",
      x = "Date"
    ) +
    theme_bw() +
    facet_wrap(~site, scales = "free_y")
```


```{r}
dens_plot <- 
  make_plot(df_summary, y = DENS_mean, ymin = DENS_lower, ymax = DENS_upper) + 
  labs(y = "Stem Density")

shift_legend(dens_plot) |> plot()
```

```{r}
npp_plot <- 
  make_plot(df_summary, y = NPP_PFT_mean, ymin = NPP_PFT_lower, ymax = NPP_PFT_upper) + 
  labs(y = "Net Primary Productivity")

shift_legend(npp_plot) |> plot()
```

```{r}
library(ggforce)
agb_plot + facet_zoom(y = pft_name == "SetariaWT")
```


```{r}
library(patchwork)


agb_plot /
  (npp_plot | dens_plot) + 
  plot_layout(guides = "collect") + plot_annotation(tag_level = "A")
```

